---
title: "SNAP retailers during the Great Recession (2008-2017): Changes in and across metro areas"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    navbar:
      - { title: "Source code", href: "https://github.com/jshannon75/snap_retailers_2008_2017/shiny/snapretailexplorer.Rmd", align: right }
      - { title: "Project site", href: "https://github.com/jshannon75/snap_retailers_2008_2017", align: right}
    
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(tidyverse)
library(leaflet)
library(sf)
library(shinyWidgets)
library(rgdal)

msa5<-read_csv("msa5_ranks_2018_07_25.csv") %>%
  arrange(msa_short)

years<-c(2008,2009,2010,2011,2012,2013,2014,2015,2016,2017)

```

Map by metro area
=====================================  

Column {.sidebar data-width=400}
-----------------------------------------------------------------------

### Data selection

This application allows you to explore the database of SNAP authorized retailers from 2008-2017 in 45 metropolitan areas, identified by their largest city.

Use the tools below to select the city, types of retailers, and years to include. Click on a point to see more information on that retailer.

```{r}
selectInput("msa","MSA selection",
            choices=msa5$msa_short)

pickerInput(
  inputId = "years2", 
  label = "Select the year(s) retailers were SNAP authorized", 
  choices = years, 
  selected=max(years),
  options = list(
    `actions-box` = TRUE, 
    size = 10,
    `selected-text-format` = "count > 4"
  ), 
  multiple = TRUE
)

checkboxGroupInput("storegroup", 
                    ("Retailer types"), 
                    choices = list("Supermarkets/supercenters" = "Supermarket", 
                                   "Groceries" = "Grocer", 
                                   "Convenience/dollar stores" = "Small retail",
                                   "Specialty retailers"="Specialty",
                                   "Local foods"="Local foods"),
                              selected = c("Supermarket","Grocer", "Small retail","Specialty","Local foods"))

HTML("If no stores are visible, click on the button below")

actionButton("loadstores1","Load stores")

HTML("<br>Download a spreadsheet of selected retailers<br>")

downloadButton('downloadData','Download Result Set')

downloadHandler(filename = function() {
     paste("SNAPretailers_",input$msa,"_",Sys.Date(),".csv",sep="")
   },
     content = function(file) {
     write.csv(msadata_df(), file, row.names = FALSE)
   }
)

HTML("<br>Data source: USDA Benefits Redemption Division<br>")


```

Column {data-height=600}
-----------------------------------------------------------------------

```{r}
#Load store data from MSA selection
msadata_reactive<-reactive({
  fips<-as.character(msa5 %>%
    dplyr::filter(msa_short==input$msa) %>%
    mutate(fips=paste("G",GEOID,sep="")) %>%
    select(fips))

  data<-st_read("storepoints_msa5.gpkg",layer=fips,stringsAsFactors=FALSE) %>%
    mutate(lat=X,long=Y,
      #Define the popup
      addr_full=if_else(is.na(addr_add),
                         paste("<br><b>",addr_num," ",addr_st," ",sep=""),
                         paste("<br><b>",addr_num," ",addr_st," ",addr_add,sep="")),
      year_range=case_when(year_first==year_last~as.character(year_first),
                           year_first!=year_last~paste(year_first,year_last,sep="-")),
      content=paste("<strong><size=1>",store_name,"</strong></size>", 
                       addr_full,
                       "<br>",city,", ",state,"  ",zip5,"</b>",
                       "<br><br><b>USDA store classification:</b> ",store_type,
                       "<br><b>Group category: </b>",store_group,
                       "<br><b>Years authorized for SNAP*: </b>",year_range,
                       "<br>*2008 and 2017 are the first and last years data are available.",sep=""))
  })

#Subset data based on store type and year
msadata_select<-reactive({
  msadata_reactive() %>%
    filter(store_group %in% input$storegroup) %>% 
    gather(Y2008:Y2017,key="year",value="present") %>%
    mutate(year_num=as.numeric(substr(year,2,5))) %>%
    filter(year_num %in% input$years2 & present==1) %>%
    select(-year_num) %>%
    distinct() %>%
    spread(year,present,fill=0)
    #filter(year_firstinput$years[1]&year_first<=input$years[2])
})

# observeEvent(c(input$msa, input$years2, input$storegroup),{
msadata_df<-reactive({
  msadata_select() %>%
  st_set_geometry(NULL) %>%
    select(-lat,-long,-addr_full,-content)
})

#Create a map  
  buffer<-0.02

output$map<-renderLeaflet({
  long_max<-max(msadata_reactive()$X)+buffer
  latit_max<-max(msadata_reactive()$Y)+buffer
  long_min<-min(msadata_reactive()$X)-buffer
  latit_min<-min(msadata_reactive()$Y)-buffer

  leaflet(msadata_reactive()) %>%
  fitBounds(long_min,latit_min,long_max,latit_max) %>%
   addProviderTiles(providers$CartoDB.Positron) 
})

observeEvent(input$msa,{
  long_max1<-max(msadata_reactive()$X)+buffer
  latit_max1<-max(msadata_reactive()$Y)+buffer
  long_min1<-min(msadata_reactive()$X)-buffer
  latit_min1<-min(msadata_reactive()$Y)-buffer
  
  leafletProxy("map") %>%
    fitBounds(long_min1,latit_min1,long_max1,latit_max1)})

observeEvent(c(input$storegroup, input$msa, input$loadstores1,input$years2),{ 
 #storepal<-colorFactor(palette="YlOrRd",msadata_reactive()$year_range,reverse=TRUE)
  storepal<-colorFactor(palette="Dark2",msadata_reactive()$store_group,reverse=TRUE)
  
 leafletProxy("map") %>%
    clearMarkers() %>%
    clearControls() %>%
    addCircleMarkers(data=msadata_select(),
               opacity=0.4,
               color="#603a31",
               weight=2,
               fillOpacity=0.8,
               radius=4,
               fillColor=~storepal(store_group),
               popup=msadata_select()$content) %>%
   addLegend(data=msadata_reactive(),"topright",pal=storepal,
             values=~store_group,title="Store category",
             opacity=1)
})

leafletOutput("map",height=900)

```

Map by county
=====================================  

Column {.sidebar data-width=400}
-----------------------------------------------------------------------

### Data selection

This map allows you to view SNAP authorized retailers from 2008-2017 by county. Type the name of the county in the county selection search box and then select the option you want. Then click on the load data button.

Use the tools below to also choose the types of retailers and years to include. Click on a point to see more information on that retailer.

```{r}
counties<-read_csv("county_list.csv") 
county_list<-counties$fullname

selectizeInput("county","County selection",
            choices=counties$fullname,
            options = list(maxOptions = 8,selectize=TRUE),
            select="Autauga County, AL")

actionButton("loaddata2","Load data")

HTML("<br>")

pickerInput(
  inputId = "years2_1", 
  label = "Select the year(s) retailers were SNAP authorized", 
  choices = years, 
  selected=max(years),
  options = list(
    `actions-box` = TRUE, 
    size = 10,
    `selected-text-format` = "count > 4"
  ), 
  multiple = TRUE
)

checkboxGroupInput("storegroup1", 
                    ("Retailer types"), 
                    choices = list("Supermarkets/supercenters" = "Supermarket", 
                                   "Groceries" = "Grocer", 
                                   "Convenience/dollar stores" = "Small retail",
                                   "Specialty retailers"="Specialty",
                                   "Local foods"="Local foods"),
                              selected = c("Supermarket","Grocer", "Small retail","Specialty","Local foods"))

HTML("<br>Download a spreadsheet of selected retailers<br>")

downloadHandler(filename = function() {
     paste("SNAPretailers_",stcty_fips_label()$stcty_fips,"_",Sys.Date(),".csv",sep="")
   },
     content = function(file) {
     write.csv(ctydata_df(), file, row.names = FALSE)
   }
)

HTML("<br><br>Data source: USDA Benefits Redemption Division<br>")

downloadButton('downloadData','Download Result Set')

```

Column {data-height=600}
-----------------------------------------------------------------------

```{r}
#Load county boundary
cty_boundary<-reactive({
  cty_fips<-counties %>%
    filter(fullname==input$county) %>%
    mutate(cty_fips_sel=paste("G",stcty_fips,sep=""),
           st_fips=paste("G",st_fips,sep="")) %>%
    select(cty_fips_sel,st_fips)
  st_fips<-cty_fips$st_fips
  cty_select<-st_read("uscounties_st.gpkg",layer=st_fips) %>%
    filter(CTY_TXT==cty_fips$cty_fips_sel)
  cty_boundary1<-as_Spatial(cty_select)
  spTransform(cty_boundary1, CRS("+init=epsg:4326"))
})

#Load store data from MSA selection
ctydata_reactive<-reactive({
  cty_name<-input$county
  
  cty_fips<-counties %>%
    dplyr::filter(fullname==cty_name) %>%
    mutate(cty_fips_sel=paste("G",st_fips,sep="")) %>%
    select(cty_fips_sel)

  data<-st_read("storepoints_state.gpkg",layer=cty_fips$cty_fips_sel,stringsAsFactors=FALSE) %>%
    mutate(lat=X,long=Y,
      addr_full=if_else(is.na(addr_add),
                         paste("<br><b>",addr_num," ",addr_st," ",sep=""),
                         paste("<br><b>",addr_num," ",addr_st," ",addr_add,sep="")),
      year_range=case_when(year_first==year_last~as.character(year_first),
                           year_first!=year_last~paste(year_first,year_last,sep="-")),
      content=paste("<strong><size=1>",store_name,"</strong></size>",
                       addr_full,
                       "<br>",city,", ",state,"  ",zip5,"</b>",
                       "<br><br><b>USDA store classification:</b> ",store_type,
                       "<br><b>Group category: </b>",store_group,
                       "<br><b>Years authorized for SNAP*: </b>",year_range,
                       "<br>*2008 and 2017 are the first and last years data are available.",sep=""))
  data
  })

#Subset data based on store type and year

#Create downloadable data

ctydata_df<-reactive({
  ctydata_select() %>%
  st_set_geometry(NULL) %>%
  select(-lat,-long,-addr_full,-content)
  })

#Subset data based on county name, store type, and year
stcty_fips_label<-reactive({counties %>%
    dplyr::filter(fullname==input$county) %>%
    mutate(stcty_fips=paste("G",stcty_fips,sep="")) %>%
    select(stcty_fips)})

 ctydata_select<-reactive({
  cty_name<-input$county
   
  stcty_fips<-counties %>%
    dplyr::filter(fullname==cty_name) %>%
    mutate(stcty_fips=paste("G",stcty_fips,sep="")) %>%
    select(stcty_fips)
  
  #Create single variable based on input year type
  yeara<-if_else(input$year_type=="By first year in data",input$years1[1],input$years1a)
  yearb<-if_else(input$year_type=="By first year in data",input$years1[2],input$years1a)
  
  ctydata_reactive() %>%
    filter(cty_fips==stcty_fips$stcty_fips) %>%
    filter(store_group %in% input$storegroup1) %>%
    gather(Y2008:Y2017,key="year",value="present") %>%
    mutate(year_num=as.numeric(substr(year,2,5))) %>%
    filter(year_num %in% input$years2_1 & present==1) %>%
    select(-year_num) %>%
    distinct() %>%
    spread(year,present,fill=0)
})

#Create a map
 buffer<-0.02

#Initial view is Autauga County, AL
output$map1<-renderLeaflet({
  leaflet() %>%
   setView(lng=-86.651654,lat=32.543148,zoom=10) %>%
   addTiles(urlTemplate="https://api.mapbox.com/styles/v1/shann039/cjdyoczre1lt02rmw9rhw2vps/tiles/256/{z}/{x}/{y}?access_token=pk.eyJ1Ijoic2hhbm4wMzkiLCJhIjoiY2prenRsMXl1MHh2czNxbnNmN3R0MGI0aSJ9.wPnB1BQ_AjhruRrlXJ6tiA")
   #addProviderTiles(providers$CartoDB.Positron) 
})

observeEvent(input$loaddata2,{
  long_max<-max(ctydata_select()$X)+buffer
  latit_max<-max(ctydata_select()$Y)+buffer
  long_min<-min(ctydata_select()$X)-buffer
  latit_min<-min(ctydata_select()$Y)-buffer

  leafletProxy("map1") %>%
    fitBounds(long_min,latit_min,long_max,latit_max)})

observeEvent(c(input$storegroup1, input$years2_1,input$loaddata2),{
  storepal<-colorFactor(palette="Dark2",ctydata_reactive()$store_group,reverse=TRUE)

 leafletProxy("map1") %>%
    clearMarkers() %>%
    clearControls() %>%
    addPolygons(data=cty_boundary(),
                fillOpacity=0,
                opacity=0.7,
                stroke=TRUE,
                color="black") %>%
    addCircleMarkers(data=ctydata_select(),
               opacity=0.4,
               color="#603a31",
               weight=2,
               fillOpacity=0.8,
               radius=4,
               fillColor=~storepal(store_group),
               popup=ctydata_select()$content) %>%
   addLegend(data=ctydata_reactive(),"topright",pal=storepal,
             values=~store_group,title="Store category",
             opacity=1)
})

 leafletOutput("map1",height=900)

```

About this map
=====================================

This map shows SNAP autorized retailers from a database of approximately 450,000 total records from 2008 through 2017. These records were obtained through a request to USDA's Benefit Redemptions Division and linked across years by name and address. The full dataset includes all authorized retailers, but this site maps the five largest metropolitan statistical areas (MSAs) in each of the nine census regions and each county in the US. Each MSA is identified by its largest city in the dropdown menu. 

The number of SNAP retailers increased significantly in the years following the Great Recession. The control tools allow users to select stores based on when they first appear in the dataset. Stores appearing in 2008 may have been present before this period, and those with a last year of 2017 may still be open, as these are the current start and end points of the data. Using the dropdown selection tool allows users to identify stores that became authorized to redeem SNAP benefits during this period. These groups are based on USDA's store classifications. The exact USDA classifications are available in the store popup windows and the downloadable spreadsheet.

Users can download store data for any city using the button provided here. The full dataset is available on the [project website](https://github.com/jshannon75/snap_retailers_2008_2017). A fuller description of the dataset is also available in [this summary document](https://jshannon75.github.io/snap_retailers_2008_2017/overview_paper).
