---
title: "Store distance calculations"
author: "Jerry Shannon"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(purrr)
library(parallel)
```

This document provides the commands used to measure the Euclidean distance between block centroids and the five closest SNAP authorized retailers in each store group. This process can be VERY time intensive, with a single city taking 10-20 hours on desktop/laptop PC. I used Google Cloud Engine to run this analysis, which allowed me to use multicore processing to reduce this time to a half hour or less for each city. When completed, I create population weighted mean values at tract level from these data.

First, I read in national data for census block centroids (which include population in 2010) and the SNAP retailer data. To aid in the later analysis, I also create a new variable for the county FIPS code in the block data. Both sets of points are converted to spatial data using st_as_sf from the sf package, which adds a geometry column.

```{r data, message=FALSE}
blocks<-st_read("http://www.shannons.us/Files/snap/blocks_msa5name.gpkg", stringsAsFactors = FALSE) %>%
  mutate(gisjn_tct=substr(GISJOIN,1,14))
blocks_df<-blocks
st_geometry(blocks_df)<-NULL

msa_ranks<-read_csv("http://www.shannons.us/Files/snap/msa5_ranks_2018_07_25.csv")

storedata_census<-read_csv("http://www.shannons.us/Files/snap/snap_retailers_crosswalk.csv")
storedata<-read_csv("http://www.shannons.us/Files/snap/snap_retailers_usda.csv") %>%
   left_join(storedata_census)

storedata_shorter<-storedata %>% filter(store_group!="Other store")

#Shift stores from wide to long for years
stores_analysis<-storedata_shorter %>%
  gather(Y2008:Y2017,key="year",value="present") %>%
  filter(present==1)

storedata_sf<-st_as_sf(storedata_shorter,coords=c("X","Y"),crs=4326,remove=FALSE)
```

Before doing the analysis, I set a buffer size of 1 decimal degree when measuring retailer distance for each block centroid, which speeds the analysis.

```{r distance_calc, message=FALSE}
buffer<-1
```

The function below does a basic distance calculation using the XY coordinates. The closest five retailers in each category for each year are retained.

```{r state/block_subset, message=FALSE}

storeDist<-function(block_select){
  #Select specific block
  block_select<-blocks %>% filter(GISJOIN==block_select)
  msa<-block_select$msa_id
   
  #Set buffer zone, which speeds up the distance calculation below.
  blockX<-block_select$X
  blockY<-block_select$Y
  xmax<-blockX+buffer
  xmin<-blockX-buffer
  ymax<-blockY+buffer
  ymin<-blockY-buffer
  
  #Subset stores to just those in the buffer to speed up
  #Also create degree based distance and sort
  stores_analysis_year<-stores_analysis %>%
    filter(X<xmax & X>xmin & Y<ymax & Y>ymin) %>%
    mutate(deg_dist=(blockX-X)^2+(blockY-Y)^2) %>%
    select(store_group,year,deg_dist,X,Y) %>%
    group_by(store_group,year) %>%
    distinct() %>% # Remove duplicates
    top_n(-5,deg_dist) %>% #Select top 5
    mutate(GISJOIN=block_select$GISJOIN)
  
  #Remove any blocks with less than five locations
  stores_analysis_count<-stores_analysis_year %>%
    group_by(store_group,year) %>%
    mutate(count=n()) %>% 
    filter(count<5) %>%
    select(store_group)
  
  stores_analysis_year<-stores_analysis_year %>%
    anti_join(stores_analysis_count)
  
  #Calculate distances from stores to selected blocks
  stores_analysis_year_dist<-st_as_sf(stores_analysis_year,coords=c("X","Y"),crs=4326,remove=FALSE) 
  knn<-data.frame(st_distance(block_select,stores_analysis_year_dist,by_element=FALSE)) %>%
    gather(key="col",value="dist")
  storedist<-bind_cols(stores_analysis_year,knn) %>%
    select(GISJOIN,store_group,year,dist) %>%
    mutate(dist_km=round(as.numeric(dist/1000),3)) %>%
    arrange(store_group,year,dist) %>% #Arrange by block and distance
    group_by(store_group,year) %>%
    mutate(rank=c("D1","D2","D3","D4","D5"))
  storedist
}

system.time(test<-storeDist(blocktest))
```

I then create a second function that applies the distance function to stores in each msa.

```{r msa_apply, message=FALSE}
#Multicore for Google Cloud
detectCores()

msaid_name<-"38060"

msa_dist<-function(msaid_name){
  blocks_df_select<-blocks_df %>% 
    filter(msa_id==msaid_name) 
  
  #storedists<-map_df(blocks_df_select$GISJOIN,storeDist)
  storedists1<-mclapply(blocks_df_select$GISJOIN,mc.cores=64,storeDist)
  storedists<-bind_rows(storedists1)
  write_csv(storedists,paste("data/blockdist_",msaid_name,".csv",sep=""))
  storedists
}

##Map blocks in all MSAs
#system.time(storedist_block<-map_df(msa$GEOID,msa_dist)

#Map block in single MSAs
system.time(storedist_tract_chicago<-msa_dist("16980")) 
system.time(storedist_tract_phoenix<-msa_dist("38060"))
system.time(storedist_tract_stlouis<-msa_dist("41180"))
system.time(storedist_tract_lasvegas<-msa_dist("29820"))
system.time(storedist_tract_desmoines<-msa_dist("19780")) #Smallest MSA
```